\textbf{Computation and Neural Manifolds}

### 2025-04-26 Neural-Manifolds/neural-manifolds_v0.tex
aiaiaiai

These representations can be expressed as manifolds.

As it tranverses the manifold we observe a change of phase which can be descirbed in information geometry and the former statistical mechanics as related to likelihood. (attention to the Hodge operator). If we consider that there are many paths to start at a simplicial representation and to achive a simplicial go and the path is the computation. We can say that all computations that are within the path of least ACTION constructively interferes and all other paths destructively interferes.

@@

>>>>

These representations can be expressed as manifolds (low-dimensional representations of behavioral signals in neural population activity).

As it traverses the manifold we observe a change of phase which can be described in information geometry and the former statistical mechanics as related to likelihood. (attention to the Hodge operator). If we consider that there are many paths to start at a simplicial representation and to achieve a simplicial goal, and the path is the computation. We can say that all computations that are within the path of least ACTION constructively interferes and all other paths destructively interferes.

@@

This property is observed in brain as the complex and heterogeneous activity of the underlying circuit mechanism only hints that the computation is happening at a lower dimensional abstract representation that is part of the much larger encoded information, a subset.

Further research has demonstrated properties and the geometry of this structure.

There is also evidence of the Approximate Bayesian Computation (ABC) extended to Information geometry where the computation is the active mechanism of the brain trying to fit the likelihood of the encoded model vs the generated model. Which can be described as the translation in the manifolds. The high dimensional encoded space allows for the emergence of these structures which are relative and have a relative entropy between both models.

We can see that behavior between cortex and hypocampus.

dorsomedial prefrontal cortex (dmPFC) and hypocampus for the Temporal Estimation task.

Other areas work similarly.

In other words, individual neurons although its complexity are biased to form (groups or populations or clusters) whose (activation or firing pattern) can be represented as a dynamical system trying to perform a particular operation (there are a few), and these (groups or populations or clusters) are biased to form larger (groups or populations or clusters) whose (activation or firing pattern) goal is trying to find the path of least action through percolation properties of this complex network and these mechanics can be best described by Lagrangian mechanics.

There are so many cool parts of the percolation in this case. It is bound by the geometry, it is described by the least action, the distribution can be described by an extended version of the ABC, it has nice well explored properties which can be borrowed from the Ising model, the high dimensionality has nice properties like emergency of fractals among many things.
